---
title: "R Notebook"
output: html_notebook
---

# Libraries

```{r}
library(tidyverse)
library(ggplot2)
library(scales)
library(GGally)
library(psych)
library(gridExtra)
library(cluster) 
library(factoextra)
```

# Data Import and Preparation

```{r}
df <- read_delim("C:/Users/pcx5/Desktop/MSD_FY20Q3_Site/MSD/MER_Structured_Datasets_Site_IM_FY18-21_20201113_v1_1_Rwanda.txt", 
                 "\t", 
                 escape_double = FALSE,
                 trim_ws = TRUE,
                 col_types = cols(.default = col_character(), 
                                  targets = col_double(),
                                  qtr1 = col_double(),
                                  qtr2 = col_double(),
                                  qtr3 = col_double(),
                                  qtr4 = col_double(),
                                  cumulative = col_double()
                 ) 
) %>%
  
  filter(indicator %in% c(
    "HTS_TST",
    "HTS_INDEX",
    "HTS_TST_POS",
    "HTS_RECENT"
  ))


df <- pivot_longer(df,
                   targets:cumulative,
                   names_to = "period",
                   values_to = "value")

df <- unite(df, 
            "period", 
            c("fiscal_year", "period"),
            sep = "_", 
            remove = T)

```

# Limit to recency sites only - a site with a 2020 cumulative value for HTS_TST_POS and HTS_RECENT 
Note: KP not reported under Total Numerator (separate analysis can be done for KP)

RTRI Mods
```{r}
unique(filter(df, standardizeddisaggregate == "Modality/Age/Sex/RTRI/HIVStatus" & period == "2020_cumulative")$modality)
```

Site Totals
```{r}
recent_site <- df %>% 
  filter(period == "2020_cumulative") %>%
  mutate(disagg = case_when(standardizeddisaggregate == "Total Numerator" ~ "Total Numerator",
                                       TRUE ~ otherdisaggregate)) %>%
  filter(disagg %in% c("Total Numerator",
                                  "Recent RTRI", "Long-Term RTRI",
                                  "Confirmed Recent", "Confirmed Long-Term")) %>%
  filter(psnu != "_Military Rwanda") %>%
  unite("new_indicator", c(indicator, disagg), remove = F) %>%
  group_by(psnu, orgunituid, new_indicator) %>%
  summarise(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  pivot_wider(names_from = "new_indicator", values_from = "value") %>%
  mutate(recent_site = ifelse(!is.na(`HTS_TST_POS_Total Numerator`) & !is.na(`HTS_RECENT_Total Numerator`), TRUE, FALSE)) %>%
  filter(recent_site == TRUE) %>%
  select(-recent_site) %>%
  mutate(COVERAGE = `HTS_RECENT_Total Numerator`/`HTS_TST_POS_Total Numerator`,
         YIELD = `HTS_TST_POS_Total Numerator`/`HTS_TST_Total Numerator`,
         RTRI_YIELD = `HTS_RECENT_Recent RTRI`/`HTS_TST_Total Numerator`,
         RITA_YIELD = `HTS_RECENT_Confirmed Recent`/`HTS_TST_Total Numerator`) %>%
  select(psnu, orgunituid,
         `HTS_TST_Total Numerator`, 
         `HTS_TST_POS_Total Numerator`, 
         `HTS_RECENT_Total Numerator`,
         `HTS_RECENT_Recent RTRI`, `HTS_RECENT_Long-Term RTRI`,
         `HTS_RECENT_Confirmed Recent`, `HTS_RECENT_Confirmed Long-Term`,
         COVERAGE, YIELD, RTRI_YIELD, RITA_YIELD) %>%
  mutate_if(is.numeric , replace_na, replace = 0)

sum(recent_site$`HTS_TST_Total Numerator`, na.rm = T)
sum(recent_site$`HTS_TST_POS_Total Numerator`, na.rm = T)
sum(recent_site$`HTS_RECENT_Total Numerator`, na.rm = T)
sum(recent_site$`HTS_RECENT_Recent RTRI`, na.rm = T)
sum(recent_site$`HTS_RECENT_Long-Term RTRI`, na.rm = T)

write.csv(recent_site,"RW_RecentSites1_FY20.csv", na = "", row.names = F)
```

```{r}
recent_site_list <- unique(recent_site %>% select(psnu, orgunituid))
```

Modified Index Cascade
```{r}
recent_cascade <- df %>% 
  filter(period == "2020_cumulative") %>%
  mutate(disagg = case_when(
                            standardizeddisaggregate == "1:Age/Sex/IndexCasesOffered" ~ "1 Index",
                            standardizeddisaggregate == "2:Age/Sex/IndexCasesAccepted" ~ "2 Accepted",
                            standardizeddisaggregate == "3:Age Aggregated/Sex/Contacts" ~ "3 Contacts",
                            standardizeddisaggregate == "4:Age/Sex/Result" &
                              otherdisaggregate == "Newly Identified" &
                              statushiv == "Positive" ~ "4 NewPOS",
                            standardizeddisaggregate == "4:Age/Sex/Result" &
                              otherdisaggregate == "Newly Identified" &
                              statushiv == "Negative" ~ "4 NewNEG",
                            standardizeddisaggregate == "4:Age/Sex/Result" &
                              otherdisaggregate == "Known at Entry" &
                              statushiv == "Positive" ~ "4 KnownPOS",
                            
                            standardizeddisaggregate == "Modality/Age/Sex/RTRI/HIVStatus" &
                              modality == "Index" &
                              otherdisaggregate == "Recent RTRI" ~ "Recent RTRI Index",
                            
                                                        standardizeddisaggregate == "Modality/Age/Sex/RTRI/HIVStatus" &
                              modality == "Index" &
                              otherdisaggregate == "Long-Term RTRI" ~ "Long-Term RTRI Index",
                            
                            
                                       TRUE ~ otherdisaggregate)) %>%
  filter(disagg %in% c(
                       "1 Index", "2 Accepted",
                       "3 Contacts",
                       "4 NewPOS", "4 NewNEG", "4 KnownPOS",
                       "Recent RTRI Index", "Long-Term RTRI Index")) %>%
  filter(psnu != "_Military Rwanda") %>%
  unite("new_indicator", c(indicator, disagg), remove = F) %>%
  group_by(psnu, orgunituid, new_indicator) %>%
  summarise(value = sum(value, na.rm = T)) %>%
  ungroup() %>%
  pivot_wider(names_from = "new_indicator", values_from = "value") %>%
  select(psnu, orgunituid,
         `HTS_INDEX_1 Index`,
         `HTS_INDEX_2 Accepted`,
         `HTS_INDEX_3 Contacts`,
         `HTS_INDEX_4 KnownPOS`,
         `HTS_INDEX_4 NewNEG`,
         `HTS_INDEX_4 NewPOS`, 
         `HTS_RECENT_Recent RTRI Index`,
         `HTS_RECENT_Long-Term RTRI Index`) %>%
  mutate_if(is.numeric , replace_na, replace = 0)

cascade_df <- left_join(recent_site, recent_cascade)

write.csv(cascade_df,"RW_RecentCascade1_FY20.csv", na = "", row.names = F)
```

# Exploration

## Descriptive Statistics for Rwanda FY2020 Cumulative

```{r}
summary(recent_site)
```

## By PSNU

```{r}
recent_site %>% split(.$psnu) %>% map(summary)
```


## Boxplots of Sites by PSNU

```{r}
ggplot(recent_site, aes(y = COVERAGE, x = psnu, col = psnu)) + 
  geom_boxplot() + theme_minimal() + theme(legend.position = "none", axis.title.x = element_blank())

ggplot(recent_site, aes(y = YIELD, x = psnu, col = psnu)) + 
  geom_boxplot() + theme_minimal() + theme(legend.position = "none", axis.title.x = element_blank())

ggplot(recent_site, aes(y = RTRI_YIELD, x = psnu, col = psnu)) + 
  geom_boxplot() + theme_minimal() + theme(legend.position = "none", axis.title.x = element_blank())

ggplot(recent_site, aes(y = RITA_YIELD, x = psnu, col = psnu)) + 
  geom_boxplot() + theme_minimal() + theme(legend.position = "none", axis.title.x = element_blank())
```

## Correlation Matrix

```{r}
ggpairs(select(recent_site, -orgunituid), aes(color = psnu)) + theme_minimal()
```

## Clustering

```{r}
test <- recent_site %>% column_to_rownames("orgunituid") %>% select(-c(psnu#,
                                                                       #`HTS_RECENT_Total Numerator`, COVERAGE, YIELD, RTRI_YIELD, RITA_YIELD
                                                                       ))
test <- scale(test)
unique(recent_site$psnu)
# Using the number of PSNUs to start off
km.res <- kmeans(test, centers = 3, nstart = 25)

aggregate(test, by=list(cluster=km.res$cluster), mean)
d <- cbind(test, cluster = km.res$cluster)

fviz_cluster(km.res, data = test)

test %>%
  as_tibble() %>%
  mutate(cluster = km.res$cluster,
         site = row.names(test)) %>%
  ggplot(aes(`HTS_TST_POS_Total Numerator`, `HTS_RECENT_Confirmed Recent`, color = factor(cluster), label = site)) +
  geom_text()
```

```{r}
k2 <- kmeans(test, centers = 2, nstart = 25)
k3 <- kmeans(test, centers = 3, nstart = 25)
k4 <- kmeans(test, centers = 4, nstart = 25)
k5 <- kmeans(test, centers = 5, nstart = 25)

# plots to compare
p1 <- fviz_cluster(k2, geom = "point", data = test) + ggtitle("k = 2")
p2 <- fviz_cluster(k3, geom = "point",  data = test) + ggtitle("k = 3")
p3 <- fviz_cluster(k4, geom = "point",  data = test) + ggtitle("k = 4")
p4 <- fviz_cluster(k5, geom = "point",  data = test) + ggtitle("k = 5")

grid.arrange(p1, p2, p3, p4, nrow = 2)
k3$cluster

d <- cbind(recent_site, cluster = k4$cluster)
```

```{r}
set.seed(123)

fviz_nbclust(test, kmeans, method = "wss")
```

```{r}
fviz_nbclust(test, kmeans, method = "silhouette")
```

```{r}
gap_stat <- clusGap(test, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
# Print the result
print(gap_stat, method = "firstmax")
fviz_gap_stat(gap_stat)
```


## Positives Mix of Modality

```{r}

```


# When did each site start reporting recency results?

```{r}
df %>%
  filter(psnu != "_Military Rwanda") %>% 
  filter(indicator == "HTS_RECENT" & standardizeddisaggregate == "Total Numerator") %>%
  filter(str_detect(period, "qtr")) %>%
  filter(!str_detect(period, "2021"))%>%
  filter(!is.na(value)) %>%
  group_by(psnu, orgunituid) %>%
  summarise(first_report = min(period),
            last_report = max(period)) %>%
  ungroup() %>%
  group_by(psnu, first_report) %>%
  summarise(n = n()) %>%
  ungroup()
```

```{r}
df %>%
  filter(psnu != "_Military Rwanda") %>% 
  filter(indicator == "HTS_RECENT" & standardizeddisaggregate == "Total Numerator") %>%
  filter(str_detect(period, "qtr")) %>%
  filter(!str_detect(period, "2021")) %>%
  filter(!is.na(value)) %>%
  group_by(psnu, orgunituid, period) %>%
  summarise(value = sum(value)) %>%
  ungroup() %>%
  mutate(Report = case_when(value != 0 ~ 1)) %>% 
  ggplot(aes(x = period, y = orgunituid, group = orgunituid,  col = Report)) + 
  geom_line() + 
  geom_point(shape = 15) + 
  theme_minimal() +
  labs(title = "Site Reporting Completeness Timelines") +
  theme(axis.text.y = element_text(size=5), 
        axis.text.x = element_text(angle=90),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  facet_wrap(~psnu, scales = "free_y")
```

# Is RTRI yield significantly different from yield?

```{r}
by.site.differences <- recent_site[,"YIELD"] - recent_site[,"RTRI_YIELD"]
hist(by.site.differences[,1])
qqnorm(by.site.differences[,1])
shapiro.test(by.site.differences[,1])
```

```{r}
site_yield <- pull(recent_site, YIELD)
site_rtri_yield <- pull(recent_site, RTRI_YIELD)
wilcox.test(site_yield, site_rtri_yield)
```

# Is RTRI yield significantly different than the RITA yield?

```{r}
by.site.differences <- recent_site[,"RTRI_YIELD"] - recent_site[,"RITA_YIELD"]
hist(by.site.differences[,1])
qqnorm(by.site.differences[,1])
shapiro.test(by.site.differences[,1])
```

```{r}
site_rita_yield <- pull(recent_site, RITA_YIELD)

wilcox.test(site_rtri_yield, site_rita_yield, paired = TRUE)
```

```{r}
cor.test(site_rtri_yield, site_rita_yield,  method="kendall")
```

```{r}
cor.test(site_rtri_yield, site_rita_yield,  method="spearman")
```

# Do we see any association between yield and RTRI yield? Are higher yield sites also finding more recent positives?

```{r}
cor.test(site_yield, site_rtri_yield,  method="kendall")
```

```{r}
cor.test(site_yield, site_rtri_yield,  method = "spearman")
```

## Testing by PSNU

```{r}
recent_site %>% select(-orgunituid) %>% split(.$psnu) %>% map(cor, method = "spearman")
```


```{r}
ggplot(recent_site, aes(x=YIELD, y=RTRI_YIELD)) + geom_point() + facet_wrap(~psnu)
```

```{r}
t <- recent_site %>% mutate(NEG = `HTS_TST_Total Numerator`-`HTS_TST_POS_Total Numerator`,
                       DENOM = NEG + `HTS_RECENT_Confirmed Long-Term`)
```

